# Installing the Earth Engine API

## Set Up Using the Command Prompt (shell/terminal)

*(Note: these instructions are for Windows)*

### Activate the conda virtual environment

After Anaconda is successfully downloaded, open the Command Prompt shell by:
    
- opening the start menu and search for 'Command Prompt'
- pressing the Windows key + R and search 'cmd'
- clicking the terminal tab in Rstudio
- launching "CMD.exe Prompt" within Anaconda Navigator

You should be in your root or home directory and see something like `C:\Users\[username]>` and your cursor

If you are in a different directory, enter `cd` and your root directory (`cd` means "change directory")

In the command prompt, enter

```{r}
anaconda3\condabin\activate
```

You should now see `(base)` in front of your user name. 

**Whenever you need to be in a conda environment (such as for the rest of this tutorial), be sure to check if you have** `(base)` **in front of your user name. If you don't, always enter the above command to enter the conda environment**

Test that conda is installed correctly by entering 

```{r}
conda --help
```

If the result is something like `'conda' not found` or `'conda' not recognized as a command`, then conda did not install correctly. 

### Create a conda envrionment for Earth Engine API 

It is always recommended to work in different environments and not your base environment. This is so you can work with different packages and dependences depending on your project. We want to name this environment `rgee` since we will use this environment to work with the `rgee` package. This environment can be named anything you want, as long as you know what environment the Earth Engine API is installed in, so it makes the most sense (and is standard) to name the environment after the main package you are working with.

In the command prompt, enter 

```{r}
conda create --name rgee
```

and enter `y` when prompted. 

Activate your conda `rgee` environment by entering 

```{r}
conda activate rgee
```

You should now see `(rgee)` in front of your user name

*Note: to exit the `rgee` environment, enter*

```{r}
conda deactivate
```

*and you should see `(base)` in front of your user name again.*


### Install the Earth Engine API

Install the Earth Engine API into this conda `rgee` environment by entering 

```{r}
conda install -c conda-forge earthengine-api==0.1.323
```

and type `y` when prompted.

`conda install` is the method of installing any packages into the conda environment. `-c conda-forge` is needed because `earthengine-api` is not a built-in package in python (kind of like how you might install a github package using `devtools::`). Note that we are installing version 0.1.323 of `earthengine-api`. This is not the most recent version of `earthengine-api`, but `rgee` was developped and will work best with this version.
    
Go ahead and also install the package `pandas` (a python package that handles data management, similar to `dplyr` in R). Enter

```{r}
conda install pandas
```

and type `y` when prompted.
    
This will automatically install all of `pandas` dependencies, like `numpy` (a package that handles scientific computation)

#### Install `gcloud`

We need to ensure that `gcloud` is installed to connect our Google credentials to Earth Engine API

Check that `gcloud` is installed or not by typing in the command shell

```{r}
glcoud --help
```

if the result is something like `'gcloud' not found` or `'gcloud' not recognized as a command`, then gcloud is not installed. (If you see something besides the above, then skip the rest of this sub-section)

Install the `gcloud` Command Line Interface (CLI) by following the instructions [here](https://cloud.google.com/sdk/docs/install)

After the installer finishes, select the option to open the Google Cloud Command Prompt.
    
If you closed out of the installer without selecting to open the Google Cloud Command Prompt, it's ok. Close and re-open the Windows Command Prompt (or enter `conda deactivate` twice until you are out of the conda environment and no longer see `(rgee)` or `(base)`) and enter 

```{r}
cd "AppData\Local\Google\Cloud SDK"
```

You are now in this folder directory. Enter

```{r}
cloud_env.bat
```

and now you are in the Google Cloud CLI

You will be prompted to log in to your Google account. Type `y` and a browser window will automatically open. Log in to the Google account that your Earth Engine account is connected to. Return to the command prompt and select the correct cloud project that you are working in.

Close and re-open the Command Prompt shell, re-activate conda and the conda `rgee` environment 

(you should see `(rgee)` in front of your user name)

### Connect your Earth Engine credentials to the API

5. In the command shell, type

```{r}
earthengine authenticate
```

A browser will automatically open. Log in to the Google account connected to your Earth Engine account and allow access to the Google Earth Engine Authenticator

\

#### Check that the API was connected correctly

In the command shell, type

```{r}
python
```

This will begin a python interpreter in your conda ee environment, allowing you to code in python in the command shell

Enter

```{r}
import ee
ee.Initialize()
print(ee.Image('USGS/SRTMGL1_003').getInfo())
```

If the command line prints out metadata of a DEM data set and looks something like 

```{r}
{'type': 'Image', 'bands': [{'id': 'elevation', 'data_type': {'type': 'PixelType', 'precision': 'int', 'min': -3
```

then you have successfully connected the API to your Earth Engine account

To exit the python interpreter, enter Ctr + Z (It will look like `^Z`)



## Set Up Using Anaconda Navigator